Titeln på avhandlingens språk

### 2022-11-06 implications_optimization.tex
working on sampling

In the following, we will show the effects of our optimization approach combined with our timing models compared to \textcite{Wang2019Towards}'s state-of-the-art approach.
\todo[inline]{Sounds clunky}
For this we implement these sampling schemes in Python and run a number of simulations with them.
The first scheme uses our approach and \cref{eq:epsilon_terminal,eq:optimization:sampling} to determine the optimum sampling instants and the values for \( \alpha \) and \( \beta \) at each step.
It includes an embedded timing model to provide updated estimates of \( \sigma \) at every step as well, allowing it to adapt to the state of the system.

@@

\begin{table*}
    \centering
    \caption{Experimental parameters}\label{tab:params}
    \begin{tabular}{lrl}
        \toprule
        Parameter & Value & Clarification \\
        \midrule
        Number of task steps & \num{100} & \\
        Repetitions & \num{100} & \\
        \acp{RTT} & \SIlist{0.3375;0.675;1.25;2.5;5.0}{\second} & \\
        \( t_\text{p} \) & \SI{300}{\milli\second} & Processing delay \\
        \( t_\text{c} \) & \( \text{\ac{RTT}} - t_\text{p} \) & Communication delay \\
        \( w_0 \) & \SI{1.0}{\second} & \\
        \( R_\text{min} \) & \SI{0.5}{\hertz} & Derived as \( R_\text{min} = {(2 w_0)}^{-1} \) \\
        \( \varphi \) & 3.0 & Scaling factor, \textcite{Wang2019Towards}\\
        \( P_0 \) & \SI{15}{\milli\watt} & Idle power \\
        \( P_\text{c} \) & \SI{45}{\milli\watt} & Communication power \\
        \bottomrule
    \end{tabular}
\end{table*}

We proceed to set up an experiment where these sampling approaches are deployed on identical, simulated, tasks.
Experimental parameters are detailed in \cref{tab:params}.
The execution times for each step are generated by a realistic model without any distribution fitting on the data.

 (this timing model is separate from the internal timing models inside the sampling schemes).
We set \( R_\text{min} \) to \SI{0.5}{\hertz} and \( \alpha \) to \num{3.0}.
The \ac{RTT} of the system is set to one of \SIlist[list-final-separator={, or}]{150;300;600}{\milli\second}, constant for all steps of the task.
For each combination of sampling scheme configuration (original adaptive by \textcite{Wang2019Towards}, adaptive with \edgedroid{} model, greedy) and \ac{RTT}, we run \num{100} repetitions of the task for good statistical significance.

\begin{figure*}
    \centering
    \begin{subtable}[t]{.25\textwidth}
        \centering
        \caption{Parameters used for \textcite{Wang2019Towards}'s adaptive sampling scheme.}
        \medskip
        \begin{tabular}{lr}
            \toprule
            \( R_\text{min} \)  & \SI{0.5}{\hertz} \\
            \( \alpha \)        & \num{3.0} \\
            \ac{RTT}            & \SIlist{150;300;600}{\milli\second} \\
            \bottomrule
        \end{tabular}
    \end{subtable}%
    \hspace{.05\textwidth}%
    \begin{subtable}[t]{.65\textwidth}
        \centering
        \caption{Mean number of captured samples per sampling scheme at the three different \acp{RTT}.}
        \medskip
        \begin{tabular}{llrrr}
            \toprule 
            Timing model neuro. & Sampling scheme & \SI{150}{\milli\second} & \SI{300}{\milli\second} & \SI{600}{\milli\second} \\
           \midrule
           \multirow[c]{2}{*}{Low} & Greedy & \num{36.53} & \num{18.88} & \num{10.52} \\
            & Adaptive Gaussian fit & \num{19.97} & \num{10.43} & \num{6.39} \\
           \multirow[c]{2}{*}{High} & Greedy & \num{37.05} & \num{19.04} & \num{10.89} \\
            & Adaptive Gaussian fit & \num{20.29} & \num{10.79} & \num{6.67} \\
            \bottomrule
        \end{tabular}
    \end{subtable}\\
    \bigskip
    \begin{subfigure}[]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/diff_sampling_junjue_both.png}
        \caption{%
            Percentage reduction in mean number of captured samples per step (higher is better).
            Error bars indicate the \SI{95}{\percent} \acp{CI}, calculated using a two-sided T-test.
            \todo[inline]{%
                % Vishnu: minimize number of samples.
                I don't like the two plots, I think the message gets diluted and confusing.
                I suggest only keeping one ``ground truth'', high neuroticism.
                Otherwise, every plot and table needs to be duplicated.
            }
        }
    \end{subfigure}
    \caption{%
        Results of parameterizing \textcite{Wang2019Towards}'s adaptive sampling scheme using the realistic timing model.
        \emph{ED2} indicates a sampling scheme using the \edgedroid{} model for \ac{CDF} estimation.
        \emph{HN} and \emph{LN} indicate high or low neuroticism, respectively, and \emph{ExG fit} indicates a timing model which fits an \ac{exGaussian} distribution to the execution time data for each step.
    }\label{fig:samples}
\end{figure*}

The results of this investigation are presented in \cref{fig:samples}
They clearly show the potentially dramatic effects of more accurate execution time estimation when optimizing sampling rates.
When using the realistic model for execution time estimation, the sampling schemes achieve between \SI{14}{\percent} and \textasciitilde\SI{27}{\percent} better average performance, depending on the \edgedroid{} model parameterization and the system \ac{RTT}.
\todo[inline]{Update with new ground truths.}
This is true even for configurations with low neuroticism which do not exactly match the parameterization of the external timing model driving the execution times for the steps, showing thus the importance of even less-accurate approximations.

@@

\subsubsection{Minimizing number of samples }\label{sec:aprxSol_Energy_samples}

\subsubsection{Evaluating the impact of the realistic timing model on the energy-optimal sampling scheme}

\begin{figure*}
    \centering
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/sampling_optimization.png}
        \caption{\todo[inline]{Note logarithmic yscale.}}
    \end{subfigure}\\
    \medskip
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/sampling_optimization_diff.png}
        \caption{}
    \end{subfigure}
    \caption{Sampling optimization}
\end{figure*}

>>>>

In the following, we will show the effects of our optimization approach combined with our timing models compared to \textcite{Wang2019Towards}'s state-of-the-art approach.
\todo[inline]{Sounds clunky}
For this we implement these sampling schemes in Python and run a number of simulations with them.
The first scheme uses our approach and \cref{eq:epsilon_terminal,eq:optimization:sampling} to determine the optimum sampling instants and the values for \( \alpha \) and \( \beta \) at each step.
It includes an embedded timing model (without any distribution fitting) to provide updated estimates of \( \sigma \) at every step as well, allowing it to adapt to the state of the system.
We will refer to this scheme as the \emph{sample-count-optimized aperiodic} sampling scheme.

@@

\begin{table}
    \centering
    \caption{Experimental parameters}\label{tab:params}
    \begin{tabular}{lrl}
        \toprule
        Parameter & Value & Clarification \\
        \midrule
        \# of steps & \num{100} & \\
        Repetitions & \num{100} & \\
        \acp{RTT} & \makecell[cr]{%
            \numlist[list-final-separator={, }]{0.3375;0.675;1.25},\\
            \SIlist{2.5;5.0}{\second}
        } & \\
        \( t_\text{p} \) & \SI{300}{\milli\second} & Processing delay \\
        \( t_\text{c} \) & \( \text{\ac{RTT}} - t_\text{p} \) & Communication delay \\
        \( w_0 \) & \SI{1.0}{\second} & \\
        \( R_\text{min} \) & \SI{0.5}{\hertz} & Derived as \( R_\text{min} = {(2 w_0)}^{-1} \) \\
        \( \varphi \) & 3.0 & Scaling factor, \textcite{Wang2019Towards}\\
        \( P_0 \) & \SI{15}{\milli\watt} & Idle power \\
        \( P_\text{c} \) & \SI{45}{\milli\watt} & Communication power \\
        \bottomrule
    \end{tabular}
\end{table}

We proceed to set up an experiment where these sampling approaches are deployed on identical, simulated, tasks.
Experimental parameters are detailed in \cref{tab:params}.
The execution times for each step are generated by a timing model without any distribution fitting on the data.
For each combination of sampling scheme configuration, \ac{RTT}, and execution time model neuroticism (low or high), we run \num{100} repetitions of the task for good statistical significance.
Note that the embedded timing model in our sampling-count-optimized aperiodic sampling scheme is always parameterized with a neuroticism matching the neuroticism of the external execution time model.

\begin{figure*}
    \centering
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/sampling_optimization.png}
        \caption{\todo[inline]{Note logarithmic yscale.}}
    \end{subfigure}\\
    \medskip
    \begin{subfigure}[t]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/new_model/sampling_optimization_diff.png}
        \caption{}
    \end{subfigure}
    \caption{Sampling optimization}\label{fig:optimization:samples}
\end{figure*}

The results of this investigation are presented in \cref{fig:optimization:samples}
They clearly show the potentially dramatic effects of more accurate execution time estimation when optimizing sampling rates.
When using the realistic model for execution time estimation, the sampling schemes achieve between \SI{14}{\percent} and \textasciitilde\SI{27}{\percent} better average performance, depending on the \edgedroid{} model parameterization and the system \ac{RTT}.
\todo[inline]{Update with new ground truths.}
This is true even for configurations with low neuroticism which do not exactly match the parameterization of the external timing model driving the execution times for the steps, showing thus the importance of even less-accurate approximations.

@@
